---
title: "STAT405 Bayesian Statistics Final Project"
output: pdf_document
---
# EDA
```{r, warning=FALSE}
library(ggplot2)
library(dplyr)
library(corrplot)
library(car)
suppressPackageStartupMessages(library(cmdstanr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(bayesplot))
source("simple.R")
source("simple_utils.R")
```

```{r}
house_price <- read.csv("data/house-prices-advanced-regression-techniques/train.csv")

# check MSSubClass, it should be character
house_price$MSSubClass <- as.character(house_price$MSSubClass)

# Replace NA with explicit categories
house_price$Alley[is.na(house_price$Alley)] <- "NoAlley"
house_price$GarageType[is.na(house_price$GarageType)] <- "NoGarageType"
house_price$MasVnrType[is.na(house_price$MasVnrType)] <- "NoMasVnrType"
house_price$MasVnrArea[is.na(house_price$MasVnrArea)] <- 0
house_price$Fence[is.na(house_price$Fence)] <- "NoFence"
house_price$MiscFeature[is.na(house_price$MiscFeature)] <- "NoMiscFeature"


house_price$BsmtQual[is.na(house_price$BsmtQual)] <- "NoBsmt"
house_price$BsmtCond[is.na(house_price$BsmtCond)] <- "NoBsmt"
house_price$BsmtExposure[is.na(house_price$BsmtExposure)] <- "NoBsmt"
house_price$BsmtFinType1[is.na(house_price$BsmtFinType1)] <- "NoBsmt"
house_price$BsmtFinType2[is.na(house_price$BsmtFinType2)] <- "NoBsmt"
house_price$FireplaceQu[is.na(house_price$FireplaceQu)] <- "NoFireplace"
house_price$GarageFinish[is.na(house_price$GarageFinish)] <- "NoGarage"
house_price$GarageQual[is.na(house_price$GarageQual)] <- "NoGarage"
house_price$GarageCond[is.na(house_price$GarageCond)] <- "NoGarage"
house_price$PoolQC[is.na(house_price$PoolQC)] <- "NoPool"

# Add indicator variables first
house_price$HasBsmt <- as.factor(ifelse(house_price$TotalBsmtSF > 0, 1, 0))
house_price$HasGarage <- as.factor(ifelse(house_price$GarageArea > 0, 1, 0))
house_price$HasFireplace <- as.factor(ifelse(house_price$Fireplaces > 0, 1, 0))
house_price$HasPool <- as.factor(ifelse(house_price$PoolArea > 0, 1, 0))
house_price$HasMasVnr <- as.factor(ifelse(house_price$MasVnrArea > 0, 1, 0))

# Convert to ordered factors
qual_levels <- c("Po", "Fa", "TA", "Gd", "Ex")
house_price$ExterQual <- factor(house_price$ExterQual,
                                levels = qual_levels, ordered = TRUE)
house_price$ExterCond <- factor(house_price$ExterCond,
                                levels = qual_levels, ordered = TRUE)
house_price$BsmtQual <- factor(house_price$BsmtQual,
                               levels = c("NoBsmt", "Po", "Fa", "TA", "Gd", "Ex"),
                               ordered = TRUE)
house_price$BsmtCond <- factor(house_price$BsmtCond,
                               levels = c("NoBsmt", "Po", "Fa", "TA", "Gd", "Ex"),
                               ordered = TRUE)
house_price$BsmtExposure <- factor(house_price$BsmtExposure,
                                   levels = c("NoBsmt", "No", "Mn", "Av", "Gd"),
                                   ordered = TRUE)
house_price$BsmtFinType1 <- factor(house_price$BsmtFinType1,
                                   levels = c("NoBsmt", "Unf", "LwQ", "Rec", "BLQ", "ALQ", "GLQ"),
                                   ordered = TRUE)
house_price$BsmtFinType2 <- factor(house_price$BsmtFinType2,
                                   levels = c("NoBsmt", "Unf", "LwQ", "Rec", "BLQ", "ALQ", "GLQ"),
                                   ordered = TRUE)
house_price$HeatingQC <- factor(house_price$HeatingQC,
                                levels = qual_levels, ordered = TRUE)
house_price$KitchenQual <- factor(house_price$KitchenQual,
                                  levels = qual_levels, ordered = TRUE)
house_price$Functional <- factor(house_price$Functional,
                                 levels = c("Sal", "Sev", "Maj2", "Maj1", "Mod", "Min2", "Min1", "Typ"),
                                 ordered = TRUE)
house_price$FireplaceQu <- factor(house_price$FireplaceQu,
                                  levels = c("NoFireplace", "Po", "Fa", "TA", "Gd", "Ex"),
                                  ordered = TRUE)
house_price$GarageFinish <- factor(house_price$GarageFinish,
                                   levels = c("NoGarage", "Unf", "RFn", "Fin"),
                                   ordered = TRUE)
house_price$GarageQual <- factor(house_price$GarageQual,
                                 levels = c("NoGarage", "Po", "Fa", "TA", "Gd", "Ex"),
                                 ordered = TRUE)
house_price$GarageCond <- factor(house_price$GarageCond,
                                 levels = c("NoGarage", "Po", "Fa", "TA", "Gd", "Ex"),
                                 ordered = TRUE)
house_price$PoolQC <- factor(house_price$PoolQC,
                             levels = c("NoPool", "Fa", "TA", "Gd", "Ex"),
                             ordered = TRUE)
house_price$LotShape <- factor(house_price$LotShape,
                               levels = c("IR3", "IR2", "IR1", "Reg"),
                               ordered = TRUE)
house_price$LandSlope <- factor(house_price$LandSlope,
                                levels = c("Sev", "Mod", "Gtl"),
                                ordered = TRUE)
house_price$PavedDrive <- factor(house_price$PavedDrive,
                                 levels = c("N", "P", "Y"),
                                 ordered = TRUE)

# Convert ordered factors to numeric scores
ordinal_to_numeric <- function(x) as.numeric(x) - 1

ord_vars <- c(
  "ExterQual", "ExterCond",
  "BsmtQual", "BsmtCond", "BsmtExposure", "BsmtFinType1", "BsmtFinType2",
  "HeatingQC", "KitchenQual", "Functional", "FireplaceQu",
  "GarageFinish", "GarageQual", "GarageCond",
  "PoolQC", "LotShape", "LandSlope", "PavedDrive"
)
house_price[ord_vars] <- lapply(house_price[ord_vars], ordinal_to_numeric)
house_price <- na.omit(house_price)

cat_vars <- names(house_price)[sapply(house_price, function(x) is.factor(x) || is.character(x))]
for (v in cat_vars) {
  cat("\n----------------------\n")
  cat("Variable:", v, "\n")
  print(table(house_price[[v]], useNA = "ifany"))
}
```


```{r}
# month as time seires
house_price$month_sin <- sin(2 * pi * house_price$MoSold / 12)
house_price$month_cos <- cos(2 * pi * house_price$MoSold / 12)
# 
house_price$YearSinceRemodel <- house_price$YrSold - house_price$YearRemodAdd
house_price$YearSinceBuild <- house_price$YrSold - house_price$YearBuilt

house_price <- house_price |>
  select(-Id, -MoSold, -YrSold, -YearRemodAdd, -YearBuilt, 
         # only one level
         -HasGarage, -Utilities,
         # too sparse
         -Street, -Condition2, -RoofMatl, -Heating, -HasPool, -MiscFeature)

# merge some level
house_price$Condition1 <- dplyr::case_when(
  house_price$Condition1 == "Norm" ~ "Norm",
  house_price$Condition1 %in% c("Artery", "Feedr") ~ "Road",
  house_price$Condition1 %in% c("RRAn", "RRAe", "RRNe", "RRNn") ~ "Railroad",
  house_price$Condition1 %in% c("PosA", "PosN") ~ "PositiveFeature",
  TRUE ~ "Other"
)
house_price$Condition1 <- factor(house_price$Condition1)
house_price$RoofStyle <- dplyr::case_when(
  house_price$RoofStyle %in% c("Gable", "Hip") ~ house_price$RoofStyle,
  TRUE ~ "Other"
)
house_price$RoofStyle <- factor(house_price$RoofStyle)
house_price$Exterior1st <- dplyr::case_when(
  house_price$Exterior1st %in% c("VinylSd", "MetalSd", "HdBoard", "Wd Sdng", "Plywood", "CemntBd") ~ house_price$Exterior1st,
  house_price$Exterior1st %in% c("BrkFace", "Stone", "Stucco", "WdShing", "AsbShng") ~ house_price$Exterior1st,
  TRUE ~ "Other"
)
house_price$Exterior1st <- factor(house_price$Exterior1st)
house_price$Exterior2nd <- dplyr::case_when(
  house_price$Exterior2nd %in% c("VinylSd", "MetalSd", "HdBoard", "Wd Sdng", "Plywood", "CmentBd", "Wd Shng") ~ house_price$Exterior2nd,
  house_price$Exterior2nd %in% c("BrkFace", "Stucco", "AsbShng") ~ house_price$Exterior2nd,
  TRUE ~ "Other"
)
house_price$Exterior2nd <- factor(house_price$Exterior2nd)
house_price$Electrical <- dplyr::case_when(
  house_price$Electrical == "SBrkr" ~ "SBrkr",
  house_price$Electrical %in% c("FuseA", "FuseF") ~ "Fuse",
  house_price$Electrical %in% c("FuseP", "Mix") ~ "PoorOrMixed",
  TRUE ~ "Other"
)
house_price$Electrical <- factor(house_price$Electrical)
house_price$GarageType <- dplyr::case_when(
  house_price$GarageType == "Attchd" ~ "Attached",
  house_price$GarageType == "Detchd" ~ "Detached",
  house_price$GarageType == "BuiltIn" ~ "BuiltIn",
  house_price$GarageType %in% c("Basment", "CarPort", "2Types") ~ "OtherGarage",
  TRUE ~ "OtherGarage"
)
house_price$GarageType <- factor(house_price$GarageType)
house_price$SaleType <- dplyr::case_when(
  house_price$SaleType == "WD" ~ "WarrantyDeed",
  house_price$SaleType == "New" ~ "NewHome",
  house_price$SaleType == "COD" ~ "CourtOrEstate",
  house_price$SaleType %in% c("CWD", "Con", "ConLD", "ConLI", "ConLw", "Oth") ~ "OtherSaleType",
  TRUE ~ "OtherSaleType"
)
house_price$SaleType <- factor(house_price$SaleType)
house_price$SaleCondition <- dplyr::case_when(
  house_price$SaleCondition == "Normal" ~ "Normal",
  house_price$SaleCondition == "Abnorml" ~ "Abnormal",
  house_price$SaleCondition == "Partial" ~ "Partial",
  house_price$SaleCondition %in% c("AdjLand", "Alloca", "Family") ~ "OtherCondition",
  TRUE ~ "OtherCondition"
)
house_price$SaleCondition <- factor(house_price$SaleCondition)

cat_vars <- names(house_price)[sapply(house_price, function(x) is.factor(x) || is.character(x))]
for (v in cat_vars) {
  cat("\n----------------------\n")
  cat("Variable:", v, "\n")
  print(table(house_price[[v]], useNA = "ifany"))
}
```

```{r}
# check distribution of Y
ggplot(house_price, aes(x = SalePrice)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "white") +
  labs(title = "Histogram of House Price",
       x = "Sale Price",
       y = "Frequency")

# Right skew -> log(SalePrice)
house_price$LogSalePrice = log(house_price$SalePrice)
ggplot(house_price, aes(x = LogSalePrice)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "white") +
  labs(title = "Histogram of House Price",
       x = "Sale Price",
       y = "Frequency")
house_price <- house_price |> select(-SalePrice)
```

```{r}
numeric_vars <- house_price[, sapply(house_price, is.numeric)]
cor_matrix <- cor(numeric_vars, use = "pairwise.complete.obs")
cor_with_log_price_abs <- abs(cor_matrix[, "LogSalePrice"])
cor_with_log_price_abs <- sort(
  cor_with_log_price_abs[names(cor_with_log_price_abs) != "LogSalePrice"],
  decreasing = TRUE
)
cor_with_log_price_abs
drop_vars <- names(cor_with_log_price_abs[cor_with_log_price_abs < 0.3])
house_price <- house_price |> select(-all_of(drop_vars))
```

```{r}
# drop high cor features
numeric_vars2 <- house_price[, sapply(house_price, is.numeric)]
numeric_vars2 <- numeric_vars2[, colnames(numeric_vars2) != "LogSalePrice"]
selected_cor_matrix <- cor(numeric_vars2, use = "pairwise.complete.obs")
high_cor_pairs <- which(
  abs(selected_cor_matrix) > 0.7 & upper.tri(selected_cor_matrix),
  arr.ind = TRUE
)
high_cor_df <- data.frame(
  var1 = rownames(selected_cor_matrix)[high_cor_pairs[, 1]],
  var2 = colnames(selected_cor_matrix)[high_cor_pairs[, 2]],
  cor  = selected_cor_matrix[high_cor_pairs]
)
high_cor_df
```


```{r}
drop_vars_highcor <- c()

for (i in 1:nrow(high_cor_df)) {
  v1 <- high_cor_df$var1[i]
  v2 <- high_cor_df$var2[i]

  c1 <- cor_with_log_price_abs[v1]
  c2 <- cor_with_log_price_abs[v2]

  if (c1 < c2) {
    drop_vars_highcor <- c(drop_vars_highcor, v1)
  } else {
    drop_vars_highcor <- c(drop_vars_highcor, v2)
  }
}

drop_vars_highcor <- unique(drop_vars_highcor)
house_price <- house_price |> select(-all_of(drop_vars_highcor))
```

```{r}
# check categorical features using ANOVA
# convert character to factor
house_price[] <- lapply(house_price, function(x) {
  if (is.character(x)) as.factor(x) else x
})
# select factor variables only
cat_features <- house_price[, sapply(house_price, is.factor)]

cat_features <- cat_features[, colnames(cat_features) != "LogSalePrice", drop = FALSE]
cat_pvalues_list <- lapply(names(cat_features), function(var) {
  fit <- aov(as.formula(paste("LogSalePrice ~", var)), data = house_price)
  p_val <- summary(fit)[[1]][["Pr(>F)"]][1]
  
  level_counts <- table(house_price[[var]])
  
  data.frame(
    variable = var,
    p_value = p_val,
    n_levels = nlevels(cat_features[[var]]),
    min_count = min(level_counts),
    stringsAsFactors = FALSE
  )
})
cat_pvalues <- do.call(rbind, cat_pvalues_list)
cat_pvalues <- cat_pvalues[order(cat_pvalues$p_value), ]
cat_pvalues
```

```{r}
#full anova model
full_model <- lm(LogSalePrice ~ ., data = house_price)
alias(full_model)
```
```{r}
house_price <- house_price|> select(-BldgType)
reduced_model <- lm(LogSalePrice ~ ., data = house_price)
anova(reduced_model)
```
```{r}
house_price <- house_price |>
  select(
    -GarageType,
    -Fence,
    -HasBsmt,
    -HasFireplace,
    -HasMasVnr,
    -YearSinceBuild,
    # *
    -Exterior2nd,
    -WoodDeckSF,
    -SaleType,
    # **
    -Alley,
    -RoofStyle,
    -FullBath,
    -GarageFinish,
    -OpenPorchSF
  )
full_model <- lm(LogSalePrice ~ ., data = house_price)
vif(full_model)
```



```{r}
str(house_price)
```

# Model
## native model(only consider some numeric variable)
some reason:






```{r}
vars <- c(
  "OverallQual",
  "GrLivArea",
  "GarageCars",
  "TotalBsmtSF",
  "YearSinceRemodel"
)
for (v in vars) {
  p <- ggplot(house_price, aes(x = .data[[v]])) +
    geom_histogram(bins = 30, fill = "skyblue", color = "black") +
    labs(
      title = paste("Histogram of", v),
      x = v,
      y = "Count"
    ) +
    theme_minimal()
  print(p)
}
```
```{r}
library(dplyr)
library(ggplot2)

# Since some predictors are right-skewed, we do feature transformation
house_price_naive <- house_price |>
  mutate(
    LogGrLivArea = log(GrLivArea),
  )

vars <- c(
  "OverallQual",
  "LogGrLivArea",
  "GarageCars",
  "TotalBsmtSF",
  "YearSinceRemodel"
)

for (v in vars) {
  p <- ggplot(house_price_naive, aes(x = .data[[v]], y = LogSalePrice)) +
    geom_point(alpha = 0.4) +
    geom_smooth(method = "lm", se = TRUE) +
    labs(
      title = paste("LogSalePrice vs", v),
      x = v,
      y = "LogSalePrice"
    ) +
    theme_minimal()
  
  print(p)
}

```

\[
y_i=\log(\text{SalePrice}_i), \qquad i=1,\dots,n
\]
\[
y_i \sim \mathcal{N}(\mu_i,\sigma^2)
\]
\[
\mu_i
=
\beta_0
+ \beta_1 \text{OverallQual}_i
+ \beta_2 \text{LogGrLivArea}_i
+ \beta_3 \text{GarageCars}_i
+ \beta_4 \text{TotalBsmtSF}_i
+ \beta_5 \text{YearSinceRemodel}_i
\]
\[
\text{LogGrLivArea}_i=\log(\text{GrLivArea}_i)
\]
\[
\beta_0 \sim \mathcal{N}(8.8,0.05^2), \qquad
\beta_1 \sim \mathcal{N}(0.10,0.03^2), \qquad
\beta_2 \sim \mathcal{N}(0.25,0.08^2),
\]
\[
\beta_3 \sim \mathcal{N}(0.06,0.03^2), \qquad
\beta_4 \sim \mathcal{N}(0.00010,0.00003^2), \qquad
\beta_5 \sim \mathcal{N}(-0.003,0.004^2)
\]
\[
\sigma \sim \text{half-Normal}(0,0.10^2)
\]

```{r}
# data for prior predictive check
house_price_prior <- house_price |>
  mutate(
    LogGrLivArea = log(GrLivArea),
    ) |>
  select(
    LogSalePrice,
    OverallQual,
    LogGrLivArea,
    GarageCars,
    TotalBsmtSF,
    YearSinceRemodel
  )

# draw one set of priors
draw_prior_params <- function() {
  list(
    beta0 = simulate(Norm(8.8, 0.05)),
    beta1 = simulate(Norm(0.10, 0.03)),        # OverallQual
    beta2 = simulate(Norm(0.25, 0.08)),        # LogGrLivArea
    beta3 = simulate(Norm(0.06, 0.03)),        # GarageCars
    beta4 = simulate(Norm(0.00010, 0.00003)),  # TotalBsmtSF
    beta5 = simulate(Norm(-0.003, 0.004)),     # YearSinceRemodel
    sigma_y  = abs(simulate(Norm(0, 0.10)))
  )
}
plot_prior_lines_one_var <- function(df, xvar, n_draws = 40) {
  xvals <- df[[xvar]]
  yvals <- df$LogSalePrice
  
  xs <- seq(
    min(xvals, na.rm = TRUE),
    max(xvals, na.rm = TRUE),
    length.out = 100
  )
  
  oq0 <- median(df$OverallQual, na.rm = TRUE)
  lg0 <- median(df$LogGrLivArea, na.rm = TRUE)
  gc0 <- median(df$GarageCars, na.rm = TRUE)
  bs0 <- median(df$TotalBsmtSF, na.rm = TRUE)
  yr0 <- median(df$YearSinceRemodel, na.rm = TRUE)
  
  plot(
    xvals, yvals,
    pch = 16,
    cex = 0.45,
    col = rgb(0, 0, 0, 0.10),
    xlab = xvar,
    ylab = "log(SalePrice)",
    main = paste("Prior lines for", xvar),
    ylim = c(10.5, 13.8)
  )
  
  for (i in 1:n_draws) {
    par_draw <- draw_prior_params()
    
    OverallQual_x     <- rep(oq0, length(xs))
    LogGrLivArea_x    <- rep(lg0, length(xs))
    GarageCars_x      <- rep(gc0, length(xs))
    TotalBsmtSF_x     <- rep(bs0, length(xs))
    YearSinceRemodel_x <- rep(yr0, length(xs))
    
    if (xvar == "OverallQual") {
      OverallQual_x <- xs
    }
    if (xvar == "LogGrLivArea") {
      LogGrLivArea_x <- xs
    }
    if (xvar == "GarageCars") {
      GarageCars_x <- xs
    }
    if (xvar == "TotalBsmtSF") {
      TotalBsmtSF_x <- xs
    }
    if (xvar == "YearSinceRemodel") {
      YearSinceRemodel_x <- xs
    }
    
    # hold neighborhood effect at 0 for slope check
    mu_line <- par_draw$beta0 +
      par_draw$beta1 * OverallQual_x +
      par_draw$beta2 * LogGrLivArea_x +
      par_draw$beta3 * GarageCars_x +
      par_draw$beta4 * TotalBsmtSF_x +
      par_draw$beta5 * YearSinceRemodel_x
    
    lines(xs, mu_line, col = rgb(0, 0, 1, 0.18))
  }
}
set.seed(1)

par(mfrow = c(3, 2), mar = c(4, 4, 3, 1))

vars_check <- c(
  "OverallQual",
  "LogGrLivArea",
  "GarageCars",
  "TotalBsmtSF",
  "YearSinceRemodel"
)

for (v in vars_check) {
  plot_prior_lines_one_var(house_price_prior, v, n_draws = 40)
}

plot.new()
par(mfrow = c(1, 1))
```
```{r}
prior_rep_once <- function(df) {
  par_draw <- draw_prior_params()
  
  mu <- par_draw$beta0 +
    par_draw$beta1 * df$OverallQual +
    par_draw$beta2 * df$LogGrLivArea +
    par_draw$beta3 * df$GarageCars +
    par_draw$beta4 * df$TotalBsmtSF +
    par_draw$beta5 * df$YearSinceRemodel 
  
  yrep <- sapply(mu, function(m) simulate(Norm(m, par_draw$sigma_y)))
  return(yrep)
}
set.seed(1)

plot(
  NULL,
  xlim = c(9, 15),
  ylim = c(0, 2.5),
  xlab = "log(SalePrice)",
  ylab = "Density",
  main = "Prior predictive densities"
)

# observed density
obs_d <- density(house_price_prior$LogSalePrice)
lines(obs_d, col = "red", lwd = 3)

# prior predictive densities
for (s in 1:30) {
  yrep <- prior_rep_once(house_price_prior)
  d <- density(yrep)
  lines(d, col = rgb(0, 0, 1, 0.15))
}

legend(
  "topright",
  legend = c("Observed", "Prior predictive"),
  col = c("red", "blue"),
  lwd = c(3, 1)
)
```
# MCMC
```{r}
house_price_stan <- house_price |>
  mutate(
    LogGrLivArea = log(GrLivArea)
  ) |>
  select(
    LogSalePrice,
    OverallQual,
    LogGrLivArea,
    GarageCars,
    TotalBsmtSF,
    YearSinceRemodel
  )

stan_data <- list(
  N = nrow(house_price_stan),
  y = house_price_stan$LogSalePrice,
  OverallQual = house_price_stan$OverallQual,
  LogGrLivArea = house_price_stan$LogGrLivArea,
  GarageCars = house_price_stan$GarageCars,
  TotalBsmtSF = house_price_stan$TotalBsmtSF,
  YearSinceRemodel = house_price_stan$YearSinceRemodel
)
```

```{r, message=FALSE}
mod <- cmdstan_model("model_naive.stan")

fit_mcmc <- mod$sample(
  data = stan_data,
  seed = 123,
  chains = 4,
  parallel_chains = 4,
  iter_warmup = 1000,
  iter_sampling = 1000,
  refresh = 500,
  adapt_delta = 0.95,
  max_treedepth = 12
)
```
```{r}
# parameter summary
summary_mcmc <- fit_mcmc$summary(
  variables = c("beta0", "beta1", "beta2", "beta3", "beta4", "beta5", "sigma")
)
summary_mcmc

# trace plot
draws_mcmc <- fit_mcmc$draws()
mcmc_trace(
  draws_mcmc,
  pars = c("beta0", "beta1", "beta2", "beta3", "beta4", "beta5", "sigma")
)

# posterior predictive check
yrep_draws <- fit_mcmc$draws("y_rep")
yrep_mat <- posterior::as_draws_matrix(yrep_draws)
yrep_mean <- colMeans(yrep_mat)

plot(
  stan_data$y, yrep_mean,
  pch = 16,
  col = rgb(0, 0, 1, 0.4),
  xlab = "Observed LogSalePrice",
  ylab = "Posterior predictive mean",
  main = "Observed vs Posterior Predictive Mean"
)
abline(0, 1, col = "red", lwd = 2)

plot(
  NULL,
  xlim = range(stan_data$y),
  ylim = c(0, 2.5),
  xlab = "log(SalePrice)",
  ylab = "Density",
  main = "Posterior predictive densities"
)
lines(density(stan_data$y), col = "red", lwd = 3)

for (s in 1:30) {
  lines(density(yrep_mat[s, ]), col = rgb(0, 0, 1, 0.15))
}

legend(
  "topright",
  legend = c("Observed", "Posterior predictive"),
  col = c("red", "blue"),
  lwd = c(3, 1)
)

# predictive accuracy
rmse_yrep_mcmc <- sqrt(mean((stan_data$y - yrep_mean)^2))
mae_yrep_mcmc  <- mean(abs(stan_data$y - yrep_mean))

rmse_yrep_mcmc
mae_yrep_mcmc
```
# VI
```{r, message=False}
fit_vi <- mod$variational(
  data = stan_data,
  seed = 123,
  output_samples = 2000,
  show_messages = FALSE,
  show_exceptions = FALSE
)
```
```{r}
summary_vi = fit_vi$summary(variables = c("beta0", "beta1", "beta2", "beta3", "beta4", "beta5", "sigma"))
summary_vi


yrep_vi_draws <- fit_vi$draws("y_rep")
yrep_vi_mat <- posterior::as_draws_matrix(yrep_vi_draws)
yrep_vi_mean <- colMeans(yrep_vi_mat)
rmse_yrep_vi <- sqrt(mean((stan_data$y - yrep_vi_mean)^2))
mae_yrep_vi  <- mean(abs(stan_data$y - yrep_vi_mean))
rmse_yrep_vi
mae_yrep_vi

```

```{r}
params <- c("beta0","beta1","beta2","beta3","beta4","beta5","sigma")

draws_mcmc <- fit_mcmc$draws(params)
draws_vi   <- fit_vi$draws(params)

mcmc_df <- as_draws_df(draws_mcmc)
vi_df   <- as_draws_df(draws_vi)

par(mfrow = c(2, 4))

for (p in params) {
  d_mcmc <- density(mcmc_df[[p]])
  d_vi   <- density(vi_df[[p]])
  
  plot(d_mcmc, main = p, xlab = p, lwd = 2)
  lines(d_vi, lwd = 2, lty = 2)
  
  legend("topright",
         legend = c("MCMC", "VI"),
         lty = c(1, 2),
         lwd = 2,
         bty = "n")
}
```
```{r}
yrep_vi_draws <- fit_vi$draws("y_rep")
yrep_vi_mat <- posterior::as_draws_matrix(yrep_vi_draws)
yrep_vi_mean <- colMeans(yrep_vi_mat)
plot(
  stan_data$y, yrep_vi_mean,
  pch = 16,
  col = rgb(0, 0, 1, 0.4),
  xlab = "Observed LogSalePrice",
  ylab = "VI posterior predictive mean",
  main = "Observed vs VI Posterior Predictive Mean"
)
abline(0, 1, col = "red", lwd = 2)
plot(
  NULL,
  xlim = range(stan_data$y),
  ylim = c(0, 2.5),
  xlab = "log(SalePrice)",
  ylab = "Density",
  main = "VI Posterior Predictive Densities"
)

lines(density(stan_data$y), col = "red", lwd = 3)

for (s in 1:30) {
  lines(density(yrep_vi_mat[s, ]), col = rgb(0, 0, 1, 0.15))
}

legend(
  "topright",
  legend = c("Observed", "Posterior predictive"),
  col = c("red", "blue"),
  lwd = c(3, 1)
)
```

```{r}
range(stan_data$y)
summary(stan_data$y)

range(yrep_vi_mat)
summary(as.vector(yrep_vi_mat))

range(yrep_vi_mean)
summary(yrep_vi_mean)
```
```{r}
mu_vi_draws <- fit_vi$draws("mu")
mu_vi_mat <- posterior::as_draws_matrix(mu_vi_draws)
mu_vi_mean <- colMeans(mu_vi_mat)

range(mu_vi_mean)
summary(mu_vi_mean)

plot(
  stan_data$y, mu_vi_mean,
  pch = 16,
  col = rgb(0, 0, 1, 0.4),
  xlab = "Observed LogSalePrice",
  ylab = "VI fitted mean",
  main = "Observed vs VI fitted mean"
)
abline(0, 1, col = "red", lwd = 2)
```
# Comparison
```{r}
compare_table <- data.frame(
  parameter = summary_mcmc$variable,
  mean_mcmc = summary_mcmc$mean,
  mean_vi   = summary_vi$mean,
  sd_mcmc   = summary_mcmc$sd,
  sd_vi     = summary_vi$sd,
  q5_mcmc   = summary_mcmc$q5,
  q5_vi     = summary_vi$q5,
  q95_mcmc  = summary_mcmc$q95,
  q95_vi    = summary_vi$q95
)
compare_table
```

总体的意思就是VI拟合很差
